Nonparametric Goodness-of-Fit Tests for Discrete Null Distributions

### 2011-09-01 rjournal/ArnoldEmerson/AE.tex
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/ks-test@77 edb9625f-4e0d-4859-8d74-9fd3b1da38cb

NOTE: NEED TO ADD SIMUALTED P-VALUES.

@@

We resolve this problem by using a combination of
two simple conservative approximations to
avoid the numerical instability. First, consider the
following inequality:
\begin{align}
\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{P} \left( \lambda_{max} \sum_{i=1}^{p} \chi^2_1 \geq x \right) \label{ineq1} \\
&= \mathbb{P} \left(\chi^2_p \geq \frac{x}{p \, \lambda_{max}} \right)
\label{ineq2}
\end{align}
The values for the weighted sum can be bounded using a simple transformation
and a chi-squared distribution of a higher degree of freedom.  Second,
consider the Markov inequality: 
\begin{align}
\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{E}
  \left[\exp\left(t\sum_{i=1}^{p}\lambda_iZ_i^2\right)\right]\exp(-tx) \\
&= \frac{\exp(-tx)}{\sqrt{\prod_{i=1}^{p}(1-2t\lambda_i)}}
\label{ineq3}
\end{align}
where the bound can be
minimized over $t \in (0,1/2\lambda_{max})$.
The upper bounds for the $p$-value given by (\ref{ineq2}) and (\ref{ineq3})
are both calculated and the smaller is used in cases where the numerical
instability of (\ref{distQ}) is a concern.

@@

data:  c(0, 1) 
D = 0.5, p-value = 0.5
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
Instead, the value of $D$ given in equation (1)
should be 0 and the associated $p$-value should be 1.  
Our revision of \code{ks.test()}
fixes this problem when the user provides a discrete distribution:
\begin{Schunk} 
\begin{Sinput}
> library(ks.test)
> dgof::ks.test(c(0, 1), ecdf(c(0, 1)))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

We demonstrate the Cram\'{e}r-von Mises tests with the same
simulated data. 
\begin{Schunk}
\begin{Sinput}
> library(cvm.test)
> cvm.test(x, ecdf(1:10))
\end{Sinput}
\begin{Soutput}
	Cramer-von Mises - W2

@@

We conclude with a toy cyclical example showing that the test is
invariant to cyclic reordering of the support.
\begin{Schunk}
\begin{Sinput}
> library(cvm.test)
> set.seed(1)
> y <- sample(1:4, 20, replace=T)
> cvm.test(y, ecdf(1:4), type='U2')
\end{Sinput}
\begin{Soutput}
	Cramer-von Mises - U2

>>>>

@@

We resolve this problem by using a combination of
two conservative approximations to
avoid the numerical instability. First, consider the
following inequality:
\begin{align}
\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) &\leq \mathbb{P} \left( \lambda_{max} \sum_{i=1}^{p} \chi^2_1 \geq x \right) \label{ineq1} \\
&= \mathbb{P} \left(\chi^2_p \geq \frac{x}{p \, \lambda_{max}} \right)
\label{ineq2}
\end{align}
The values for the weighted sum can be bounded using a simple transformation
and a chi-squared distribution of a higher degree of freedom.  Second,
consider the Markov inequality: 
\begin{align}
\mathbb{P} \left(\sum_{i=1}^{p} \lambda_i \chi^2_1 \geq x \right) & \leq \nonumber \\
\mathbb{E}\biggl[\exp\biggl( & t\sum_{i=1}^{p} 
    \lambda_iZ_i^2\biggr)\biggr] \exp(-tx) \\
& = \frac{\exp(-tx)}{\sqrt{\prod_{i=1}^{p}(1-2t\lambda_i)}}
\label{ineq3}
\end{align}
where the bound can be
minimized over $t \in (0,1/2\lambda_{max})$.
The upper bounds for the $p$-value given by (\ref{ineq2}) and (\ref{ineq3})
are both calculated and the smaller is used in cases where the numerical
instability of (\ref{distQ}) may be a concern.

@@

data:  c(0, 1) 
D = 0.5, p-value = 0.5
alternative hypothesis: two-sided 
\end{Soutput}
\end{Schunk}
Instead, the value of $D$ given in equation (1)
should be 0 and the associated $p$-value should be 1.  
Our revision of \code{ks.test()}
fixes this problem when the user provides a discrete distribution:
\begin{Schunk} 
\begin{Sinput}
> library(dgof)
> dgof::ks.test(c(0, 1), ecdf(c(0, 1)))
\end{Sinput}
\begin{Soutput}
	One-sample Kolmogorov-Smirnov test

@@

We demonstrate the Cram\'{e}r-von Mises tests with the same
simulated data. 
\begin{Schunk}
\begin{Sinput}
> cvm.test(x, ecdf(1:10))
\end{Sinput}
\begin{Soutput}
	Cramer-von Mises - W2

@@

We conclude with a toy cyclical example showing that the test is
invariant to cyclic reordering of the support.
\begin{Schunk}
\begin{Sinput}
> set.seed(1)
> y <- sample(1:4, 20, replace=T)
> cvm.test(y, ecdf(1:4), type='U2')
\end{Sinput}
\begin{Soutput}
	Cramer-von Mises - U2
